---
title: "heartRateNormalModel"
author: "Bryan Milstead"
date: "December 1, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r means, include=FALSE, echo=FALSE, cache=FALSE} 
meanA<-rnorm(1,58.8,1)
meanC<-rnorm(1,74.47,1)
sdA<-rnorm(1,4.38,.25)
sdC<-rnorm(1,9.24,.25)
```

## Background

Researchers at Pusan National University in Korea compared the cardiovascular fitness of marathon runners to a control group of non-athletes. Athletes and control group members had similar ages and weights. They found a dramatic difference in the resting heart rates (measured in beats per minute [bpm]) of the athletes compared to the control group.  Marathon athletes (N=406) had a mean resting heart rate of `r round(meanA,2)`bpm with a standard deviation of `r round(sdA,2)`bpm.  The control group (N=394) had a mean resting heart rate of `r round(meanC,2)`bpm with a standard deviation of `r round(sdC,2)`bpm.

```{r data, include=FALSE, echo=FALSE, cache=FALSE} 
athletes<-(rnorm(406,58.8,4.38))
control<-(rnorm(394,74.47,9.24))
```


```{r calcs, echo=FALSE}
Xlim<-c(30,110)
Ylim<-c(0,140)
Breaks<-seq(Xlim[1],Xlim[2],4)
par(mfrow=c(1,2))
hist(athletes,xlim=Xlim,ylim=Ylim,breaks=Breaks)
hist(control,xlim=Xlim,ylim=Ylim,breaks=Breaks)
```

## Questions

- Is it appropriate to use the normal model to compare these two distributions? Yes / No

- Explain your answer.

- A nurse practitioner examines two subjects from the above study.  The first, a marathon runner, has a resting heart rate of `r round(meanA-rnorm(1,2.5,.2)*sdA)`. The second is from the control group and has a resting heart rate of `r round(meanC-rnorm(1,1.5,.2)*sdC)`.
    - Calculate the Z scores for both subjects:  Marathoner Z=            Control Z=   
    - Which one has a lower resting heart rate compared to other members of their group (athletes vs. control)?
    - Explain what the Z score for the marathon runner means.

- A subjects chosen at random has a resting heart rate of `r round(meanA+(1.3*sdA))`bpm.  
    - Is this person more likely to be a marathon runner or a member of the control group (marathon / control)?
    - Explain your reasoning.

- A: Using the 68-95-99.7 rule what is the range of resting heart rate values you would expect for the central 68% of the marathon group.  Lower range=      ?   Upper range=       ?
- B: Using the 68-95-99.7 rule what is the range of resting heart rate values you would expect for the central 95% of the control group.  Lower range=      ?   Upper range=       ?

- A: What percent of the control group will have a heart rate between `r round(meanC-(.4*sdC))`bpm and `r round(meanC+(.6*sdC))`bpm
- B: What percent of the marathon group will have a heart rate between `r round(meanA-(.6*sdA))`bpm and `r round(meanA+(.6*sdA))`bpm



