---
title: "prism"
author: "Bryan Milstead"
date: "June 29, 2018"
output: html_document
---

### Source
[Flat out stolen](https://raw.githubusercontent.com/ropensci/prism/master/README.Rmd).  Look [here](http://ropensci.github.io/prism/) also. https://github.com/ropensci/prism

https://rstudio-pubs-static.s3.amazonaws.com/260164_978471fad4f248428a52b8ff72b774be.html

##`prism`

This package allows users to access and visualize data from the [Oregon State PRISM project](http://www.prism.oregonstate.edu/).  Data is all in the form of gridded rasters for the continental US at 3 different scales: daily, monthly and 30 year normals.  Please see their webpage for a full description of the data products, or [see their overview](http://www.prism.oregonstate.edu/documents/PRISM_datasets_aug2013.pdf).

### Quickstart
Currently the package is only available on github and installable with `devtools`
```{r start, eval=FALSE}
# library(devtools)
# install_github(repo = "prism", username = "ropensci/prism")

library(prism)
library(tidyverse)
library(raster)

options(prism.path = here::here("data_local/prism")) 
```

### Downloading data

Data is available in 3 different forms as mentioned above.  Each one has it's own function to download data. While each data type has slightly different temporal parameters, the type options are always the same.  Keep in mind these are modeled parameters, not measured.  Please see the [full description](http://www.prism.oregonstate.edu/documents/Daly2008_PhysiographicMapping_IntJnlClim.pdf) for how they are calculated.

| Parameter name| Descrption           |
|:---------------:|:-------------:|
| *tmean*      | Mean temperature |
| *tmax*      | Maximum temperature      |
| *tmin* | Minimum temperature      |
| *ppt*  | Total precipitation (Rain and snow)|
| *vpdmin* | Daily minimum vapor pressure deficit [averaged over all days in the
month - normal data only] |
| *vpdmax* |Daily maximum vapor pressure deficit [averaged over all days in
the month - normal data only] |

**Normal data**

Normals are based on the years 1981 - 2010, and can be downloaded in two resolutions, `4km` and `800m`, and a resolution must be specified.  Normals can also be downloaded for a given month, vector of months, or an average for all 30 years.

```{r get normals}
get_prism_normals(type="tmean",resolution = "4km",mon = 1:6, keepZip=F)

```

The first thing to note is that you'll need to set a local location to work with this data.  Second is the option `keepZip`.  If this is `TRUE` the zip file will remain on your machine, otherwise it will be automatically deleted.

You can also view all the data you have downloaded with a simple command `ls_prism_data()`.  By default this just gives a list of file names.  All the internal functions in the package work off of this simple list of files.
```{r listing files}
## Truncated to keep file list short
ls_prism_data()[1:10,]

```

While internal plotting functions use this, other files may want an absolute path (e.g. the `raster` package), there's a parameter `absPath` that conventiently returns the absolute path.  Alternatively you may want to see what the normal name for the product is (not the file name), and that parameter is `name`.
```{r more listing}
ls_prism_data(absPath = TRUE)[1:10,]

ls_prism_data(name = TRUE)[1:10,]

#bryan: to use these absolute paths you will need to tweak it a bit as below

i <- 1 # which row from ls_prism_data() do you want to convert to raster
a <- raster(ls_prism_data(absPath=T)[i,2])

table(is.na(a[1:27000]))

freq(a)



x <- values(a)
x <- ifelse(is.na(x), 0, 1)
b <- a
values(b) <- x
plot(b)

```

You can easily make a quick plot of your data to using the output of `ls_prism_data()`

```{r quick_plot,fig.height=5,fig.width=7}
prism_image(ls_prism_data()[1,1])

# bryan: or if you converted to raster above
plot(a)


```

**Monthly and Daily Data**

Monthly and daily data is also easily accessible. Below we'll get January data for the years 1990 to 2000. We an also grab data from June 1 to June 14 2013.

```{r get daily monthly, message=FALSE}
get_prism_monthlys(type="tmean", year = 1990:2000, mon = 1, keepZip=F)
  

get_prism_dailys(type="tmean", minDate = "2013-06-01", maxDate = "2013-06-14", keepZip=F)


ls_prism_data()
```

Note that for daily data you need to give a well formed date string in the form of "YYYY-MM-DD"

You can also visualize a single point across a set of rasters.  This procedure will take a set of rasters, create a stack, extract data at a point, and then create a ggplot2 object.

Let's get make a plot of January temperatures is Boulder between 1982 and 2014.  First we'll grab all the data from the US, and then give our function a point to get data from.  The point must be a vector in the form of longitude, latitude.

```{r plot_Boulder,fig.height=5,fig.width=7}

boulder <- c(-105.2797,40.0176)
## Get data.
get_prism_monthlys(type="tmean", year = 1982:2014, mon = 1, keepZip=F)
## We'll use regular expressions to grep through the list and get data only from the month of January

ls_prism_data()

to_slice <- grep("_[0-9]{4}[0][1]",ls_prism_data()[,1],value=T)

  # wow, this took some unpacking.  For the grep statement we are building a pattern
  # the search string is "_[0-9]{4}[0][1]"
    # this defines a seven character string
    # The first is "_"
    # characters 2 to 5 are defined by [0-9]{4} this is creates all combinations
      # of the digits [0-9] {4} 
        # this corresponds to the year
    # character 6 is fixed as the digit [0]
        # this is the first digit of the month
    # character 7 is fixed as the digit [1]
        # this is the second digit of the month
    # the end result is a group of 10000 patterns to search for most are not relevant
    # based on the get_prism_monthlys() call above the relevant patterns are
        # (198201, 198301, ... 201301, 201401) 

to_slice = grep("tmean",to_slice, value = T)
    # this limits the selected files to the ones with mean temp (tmean)
    
p <- prism_slice(boulder,to_slice) 
    # This function will plot a slice of data at a single point location 
    # from a list of prism files: uses ggplot
p + stat_smooth(method="lm",se=F) + theme_bw() + 
ggtitle("Average January temperature in Boulder, CO 1982-2014")


prism_md(to_slice, returnDate = TRUE)
```


Lastly it's easy to just load up the prism data with the raster package.  This time what we'll look at January temperature anomalies.  To do this we'll examine the difference between January 2013 and the 30 year normals for January.  Conveniently, we've already downloaded both of these files.  We just need to grab them out of our list.

```{r raster_math,fig.height=5,fig.width=7}
library(raster)
### I got these just by looking at the list output
jnorm <- ls_prism_data(absPath=T)[1,2]
j2013 <- ls_prism_data(absPath=T)[52,2]
## See that the full path is returned
jnorm
## Now we'll load the rasters.
jnorm_rast <- raster(jnorm)
j2013_rast <- raster(j2013)
## Now we can do simple subtraction to get the anomaly by subtracting 2014 from the 30 year normal map
anomCalc <- function(x, y) {
  return(x - y)
  }

anom_rast <- overlay(j2013_rast,jnorm_rast,fun = anomCalc)

plot(anom_rast)


```

The plot shows that January 2013 was warmer than the average over the last 30 years.  It also shows how easy it is to use the raster library to work with prism data.  The package provides a simple framework to work with a large number of rasters that you can easily download and vizualize or use with other data sets.


**Javascript maps**

The [leaflet package](https://cran.r-project.org/web/packages/leaflet/) allows you to easily make javascript maps using the [leaflet](http://leafletjs.com/) mapping framework using prism data.  These can easily be hosted on websites like [Rpubs](https://rpubs.com/) or your own site.  Here is a simple example of plotting the [30 year normal for annual temperature](https://rpubs.com/DistribEcology/122453). 

```{r leaflet,eval=F}
library(leaflet)
library(raster)
library(prism)

## Set default path for raster files
options(prism.path = "~/prismtmp")
get_prism_normals(type="tmean",resolution = "4km",annual =T, keepZip=F)


## Grab the data, ls_prism_data(absPath=T) will show you all the data you've downloaded
norm <- grep("tmean_30yr_normal_4kmM2_annual",ls_prism_data(absPath=T)[,2],value=T)
rast <- raster(norm)


## Create color pallette and plot
pal <- colorNumeric(c("#0000FF", "#FFFF00", "#FF0000"), values(rast),
                    na.color = "transparent")

leaflet() %>% addTiles(urlTemplate = 'http://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}') %>% 
  addRasterImage(rast,colors = pal,opacity=.65) %>% addLegend(pal = pal, values = values(rast),
                                                              title = "Deg C")

```


## Option 2
[also stolen](https://rstudio-pubs-static.s3.amazonaws.com/260164_978471fad4f248428a52b8ff72b774be.html)

* 30-yr normals Based on data from 1981-2010. Pull 30 year precipitation normals for July at a 4km grid resolution

```{r leaflet, eval=F}
#options(prism.path="/Volumes/collnell/CAT/data/prism/prism_temp")

get_prism_normals(type = 'tmean', resolution = '4km', mon = 12, keepZip = TRUE)

# ‘mon’ refers to the month of interest (July). Exchanging ‘mon’ for ‘annual = TRUE’ will download the normals for the whole year.

#Look at the downloaded prism files in your directory

ls_prism_data(name=TRUE)

```


