---
title: "RMD"
author: "Bryan Milstead"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

* add current date to yaml header (see above): "`r format(Sys.time(), '%d %B %Y')`"

relative paths

* ".." up one directory
* to go to a subdirectory "subdirectory" or "./subdirectory"
* to knit a file you may need to go up one directory first. Depends where the rmd file is.

ug/l with greek letter mu: $\mu$g/l

* no options works
```{r test} 
print('test')
```

* {r test1, include=FALSE, echo=FALSE, cache=FALSE, eval=FALSE} nothing happens of course

```{r test1, include=FALSE, echo=FALSE, cache=FALSE, eval=FALSE} 
print('test1')
```

* {r test2, include=FALSE}:no code or output

```{r test2, include=FALSE} 
print('test2')
```

* {r test3, include=TRUE}:code & output; default

```{r test3, include=TRUE} 
print('test3')
```

* {r test4, echo=FALSE}:output only

```{r test4, echo=FALSE} 
print('test4')
```

* {r test5, echo=FALSE, include=TRUE}:output only

```{r test5, echo=FALSE, include=TRUE} 
print('test5')
```

* How do you get the rcode but not the output?