---
title: "RSQLite"
author: "bryan"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(RSQLite)
```

## R Markdown

# create or connect to the database
means <- dbConnect(RSQLite::SQLite(),     
                    here::here("data_local/lake_temperatures.sqlite"))

# write table                    
dbWriteTable(means, "means", a, append = TRUE)
  print(memory.size())

# list tables
dbListTables(means)

# rename table
dbExecute(means, 'ALTER TABLE means RENAME TO tmeans_raw')

# list fields in a table
dbListFields(means, "tmeans_raw")

# index a tablee
dbExecute(means,"CREATE INDEX index_tmeans_raw_cnum ON tmeans_raw (cellnum)")
dbExecute(means,"CREATE INDEX index_tmeans_raw_year ON tmeans_raw (year)")

# submit queries
* dbGetQuery(means, "SELECT * FROM sqlite_master WHERE type = 'index'")
* dbGetQuery(means, 'SELECT Count(*) FROM nla')
* mean1 <- dbGetQuery(means, paste0('SELECT * FROM tmeans WHERE cellnum = ',
            cn, ' AND date = ', date1))
* mean2 <- dbGetQuery(means, paste0('SELECT * FROM tmeans_raw WHERE cellnum = 
            ', cn, ' AND year = ', yr1))
aves <- dbGetQuery(means, paste0('SELECT * FROM tmeans WHERE date IN 
          (20010601, 20010707, 20010930) AND
          cellnum IN (', paste(temp$cellnum, collapse=", "), ')')) 

                    
# close and save the database
  dbDisconnect(means)
  
  
  
  ```{r rsqlite, include=FALSE, echo=FALSE, cache=FALSE, eval = FALSE} 
# create or connect to the database
cc <- dbConnect(RSQLite::SQLite(),     
                    here::here("calendarOfCaptures/cc.sqlite"))

# create table captures                  
dbWriteTable(cc, "captures", caps, append = FALSE)

# create table dates                  
dbWriteTable(cc, "dates", dates, append = FALSE)

# list tables
dbListTables(cc)

# list fields in a table
dbListFields(cc, "captures")
dbListFields(cc, "dates")

# add indices
dbExecute(cc,"CREATE INDEX idx_captures_id ON captures (id)")
dbExecute(cc,"CREATE INDEX idx_captures_gr ON captures (gr)")
dbExecute(cc,"CREATE INDEX idx_captures_sp ON captures (sp)")
dbExecute(cc,"CREATE INDEX idx_captures_fjdate ON captures (fjdate)")

# close and save the database
  dbDisconnect(means)
```

 







