---
title: "STA220fall2016Grades"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gdata)
library(dplyr)
```

* read in the data (df=='grades')
* analyze grades
    - remove observations with missing values for the final (exams for 3 students still being checked and 1 did not take the exam)
    - four students have a penalty of 6.75 points on project for fabricating data.  For the analysis add 6.75 points to project grade and final grade.
* convert all score to percents
* The histogram shows the total points for the raw data (all exams and all penalties included)

```{r data, include=FALSE}
file<-'C:/Users/WMilstea/Downloads/STA220_finalGrades.xlsx'
require(gdata)
grades <- read.xls (file, sheet = 'gradeBook', header = TRUE)
grades_raw<-grades

#remove penalty for analysis
grades$project[grades$projectPoints==6.8]<-13.5
grades$totalPoints[grades$projectPoints==6.8]<-grades$totalPoints[grades$projectPoints==6.8]+6.75

#remove lines for final==NA
grades<-grades[!is.na(grades$exam2),]

grades$project<-grades$project*5
grades$exam2<-100*grades$exam2/150

names(grades)[5]<-'HW'
```

```{r analysis, echo=FALSE}
summary(grades[,4:9])
boxplot(grades[,4:9])
hist(grades_raw$totalPoints,breaks=seq(20,100,by=5))


```
