---
title: "Grades: STA220 Sec 0002 fall 2016"
author:	Bryan Milstead
date:	"23 Dec 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gdata)
library(readxl)
library(dplyr)
```

* read in the data (df=='grades') for section 002
* read in the section 1 grades (df='sec1') for comparison
* convert all scores to percents


```{r data, include=FALSE}
file<-'D:/Dropbox/220_Fall2016_Bryan/grades/STA220_finalGrades.xlsx'
excel_sheets(file)

grades<-read_excel(file, sheet = 'gradeBook')
sec1<-read_excel(file, sheet = 'sec1grades')
  


#convert project and final scores to percent
grades$project<-grades$project*5
grades$exam2<-100*grades$exam2/150

sec1$project<-sec1$project*5
sec1$exam2<-100*sec1$exam2/150

names(grades)[5]<-'HW'
names(sec1)[3]<-'HW'

#keep a copy of the grades
grades_raw<-grades
sec1_raw<-sec1
```

* calc summary stats for section 002

```{r analysis1, echo=FALSE}
summary(grades[,4:9])
```
* calc summary stats for section 001

```{r analysis1a, echo=FALSE}
summary(sec1[,c(13,3:7)])
```

* boxplots of all grades

```{r analysis2, fig.width=7, fig.height=4,echo=FALSE}
boxplot(grades[,4:9],main='sec002')
boxplot(sec1[,c(13,3:7)],main='sec001')
```

* histogram of total points 

```{r analysis3, fig.width=7, fig.height=4,echo=FALSE}
hist(grades$totalPoints,breaks=seq(20,100,by=5),main='sec002')
hist(sec1$totalPoints,breaks=seq(20,100,by=5),main='sec001')
```

```{r letterGradesFunction, include=FALSE}
#shift totalPoints up by:

letterGrades<-function(tp=grades$totalPoints,bp=0){
  a<-round(tp+bp)
  letterGrade<-rep(NA,length(a))
    letterGrade<-ifelse(a<60,'F',letterGrade)
    letterGrade<-ifelse(a<60,'F',letterGrade)
    letterGrade<-ifelse(a>=60,'D',letterGrade)
    letterGrade<-ifelse(a>=65,'D+',letterGrade)
    letterGrade<-ifelse(a>=70,'C-',letterGrade)
    letterGrade<-ifelse(a>=73,'C',letterGrade)
    letterGrade<-ifelse(a>=77,'C+',letterGrade)
    letterGrade<-ifelse(a>=80,'B-',letterGrade)
    letterGrade<-ifelse(a>=83,'B',letterGrade)
    letterGrade<-ifelse(a>=87,'B+',letterGrade)
    letterGrade<-ifelse(a>=90,'A-',letterGrade)
    letterGrade<-ifelse(a>=93,'A',letterGrade)
letterGrade<-factor(letterGrade, levels=c('A','A-','B+','B','B-','C+','C','C-','D+','D','F'))
return(letterGrade)
}


```


* totalPoints rounded to the next highest integer and converted to letter grades with the following scale:

totalPoints  | letterGrade
------------- | -------------
A  | 94-100
A- | 90-93
B+ | 87-89
B | 83-86
B- | 80-82
C+ | 77-79
C | 73-76
C- | 70-72
D+ | 65-69
D | 60-64
F | Below 60 

* compare the letter grade distribution between sections
* can also add extrapoints to each total if needed.

```{r letterGrades, echo=TRUE}

a<-table(letterGrades(grades$totalPoints,0))
b<-table(letterGrades(sec1$totalPoints,0))

a
b
barplot(a,main='sec0002')
barplot(b,main='sec0001')

# * print letterGrades by totalPoints for sec 0002

#select(arrange(grades,desc(totalPoints)),totalPoints,letterGrade)
```



```

