---
title: "vegan_package"
author: "Bryan Milstead"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(vegan)
library(MASS)
```



* There are several good vignettes and tutorials on the vegan package such as: 
    - [non-metric multidimiensional scaling](http://strata.uga.edu/software/pdf/mdsTutorial.pdf)
    - [Vegan: an introduction to ordination](https://cran.r-project.org/web/packages/vegan/vignettes/intro-vegan.pdf)
    - [Multivariate Analysis of Ecological Communities in R: vegan tutorial](http://cc.oulu.fi/~jarioksa/opetus/metodi/vegantutor.pdf)
    - [Design decisions and implementation details in vegan](https://cran.r-project.org/web/packages/vegan/vignettes/decision-vegan.pdf)
    - [Diagrams and Procedures for Partition of
Variation](https://cran.r-project.org/web/packages/vegan/vignettes/partitioning.pdf)

* From: Vegan: an introduction to ordination

```{r veganIntro, include=FALSE, echo=FALSE, eval=TRUE}

data(dune)
ord <- metaMDS(dune)
ord
plot(ord)

plot(ord, type = "n")
  points(ord, display = "sites", cex = 0.8, pch=21, col="red", bg="yellow")
  text(ord, display = "spec", cex=0.7, col="blue")
  
data(dune.env)
  attach(dune.env)
  plot(ord, disp="sites", type="n")
  ordihull(ord, Management, col=1:4, lwd=3)
  ordiellipse(ord, Management, col=1:4, kind = "ehull", lwd=3)
  ordiellipse(ord, Management, col=1:4, draw="polygon")
  ordispider(ord, Management, col=1:4, label = TRUE)
  points(ord, disp="sites", pch=21, col="red", bg="yellow", cex=1.3)
  
ord.fit <- envfit(ord ~ A1 + Management, data=dune.env, perm=999)
  ord.fit
  
plot(ord, dis="site")
  plot(ord.fit)
ordisurf(ord, A1, add=TRUE)
```

* From Multivariate Analysis of Ecological Communities in R: vegan tutorial

```{r veganTutorial, include=FALSE, echo=FALSE, eval=TRUE}
data(varespec)
  vare.dis <- vegdist(varespec)
  vare.mds0 <- isoMDS(vare.dis)
stressplot(vare.mds0, vare.dis)
ordiplot(vare.mds0, type = "t")

vare.mds <- metaMDS(varespec, trace = FALSE)
  vare.mds
  plot(vare.mds, type = "t")
  
  
data(varechem)
rankindex(scale(varechem), varespec, c("euc","man","bray","jac","kul"))




```





